---
title: "workflow for stsh"
author: "Bindoff, A."
date: "25-9-2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# reinstall TwilightFree for v 0.1.1
# devtools::install_github("ABindoff/TwilightFree")

library(readr)
library(SGAT)
library(BAStag)
library(TwilightFree)
library(lubridate)
library(dplyr)

```

Read lux data in and make Date and Light columns in correct formats  

```{r}

d <- read_delim("690__WI_FEB2015driftadj.lux", delim = "\t", skip = 19)
names(d) <- c("Date", "Light")
d$Date <- ymd_hms(dmy_hms(d$Date, tz = "UTC"))
d$Temp <- NA  # have to supply a Temp column to TwilightFree
d <- numberDays(d) # need Day_num column for some TwilightFree functions
head(d, n = 3);tail(d, n = 3)
```

Make (coarse) grid and supply colony location for a quick first-pass  

```{r}
retrieved.at <- deployed.at <- c(147.67, -43.133) # Wedge Island

grid <- makeGrid(c(40, 220), c(-65, 55),
                 cell.size = 4, mask = "sea", pacific = TRUE)
```

Determine a threshold from a (known) location and (assumed) zenith  


```{r}
thresh <- calibrate(d, as.POSIXct("2015-02-21 11:00:00"), 147.67, -43.14, zenith = 94)*1.07
```


Subset the first year of data to more easily remove non-solar light using eraseLight() (which allows you to mark bottom-left and top-right corners of a box bounding non-solar light to delete those observations), and thresholdPlot() (which plots where light crosses threshold)  

```{r}
d <- dplyr::filter(d, Date < "2015-12-30", Date > "2015-02-18")
```

```{r}
thresholdPlot(d, thresh, offset = 4)

# run the following line directly from console to erase non-solar light
# d <- eraseLight(d, thresh, offset = 4)
```

```{r, echo = FALSE}
load(file = "d_690__WI_FEB2015")
```

Find days at colony  

```{r}
colony <- findColony(d, thresh, deployed.at, q = 0.9)
missingPlot(d, threshold = thresh, fixd = colony)
```

Specify a model and fit it  

```{r, warning = FALSE}
model <- TwilightFree(d,
                      alpha = c(1, .08),
                      beta = c(1, 0.2),
                      zenith = 94, threshold = thresh, 
                      fixd = colony)
# fit the model using the forward-backward algorithm, SGAT::essie
fit <- SGAT::essie(model, grid, epsilon1=1.0E-4, epsilon2 = 1E-4)
```

It will look terrible and there will be lots of warnings, don't panic.  

```{r}
drawTracks(trip(fit))

```

We can have a look at which days went wrong specifically (highlighted), then either use `eraseLight()` and fix up obvious non-solar light and repeat the above fit, or remove the bad days entirely (shown below)  

```{r}
missingPlot(d, fit, threshold = thresh)

```

We can remove all of these days easily -  

```{r}
d <- removeMissing(d, fit)
```

(But if you'd noticed there was non-solar light on any of them, better to use `eraseLight` and run it again before resorting to the above)  

Now we can fit the track again (this time we'll provide a finer grid because hopefully this one will be a keeper)  

```{r}
grid <- makeGrid(c(40, 220), c(-65, 55),
                 cell.size = 1, mask = "sea", pacific = TRUE)

model <- TwilightFree(d,
                      alpha = c(1, .08),
                      beta = c(1, 0.2),
                      zenith = 94, threshold = thresh, 
                      fixd = colony)
# fit the model using the forward-backward algorithm, SGAT::essie
fit <- SGAT::essie(model, grid, epsilon1=1.0E-16, epsilon2 = 1E-16)

track <- trip(fit)
drawTracks(track)
```